---
title: "Custom PM Dash"
author: "Sabrina Kan"
format:
  html:
    theme: cosmo
    toc: true
    toc-location: right
    code-fold: true
---

## Palm Beach MPO Federal Performance Measures Scorecard

```{r}
library(ggplot2)

data <- data.frame(
  category = c("A", "B", "C"),
  value = c(10, 20, 15)
)

ggplot(data, aes(x = category, y = value)) +
  geom_col()
```
# 1) Safety 

## Fatalities

```{r, echo=FALSE}
library(highcharter)

fatalities <- data.frame(
  Year = c(2018, 2019, 2020, 2021, 2022, 2023, 2024),
  Number = c(183, 174, 185, 215, 226, 181, 210)
)

hchart(fatalities, "column", hcaes(x = Year, y = Number)) |>
  hc_title(text = "Total Fatalities by Year") |>
  hc_subtitle(text = "Palm Beach County") |>
  hc_yAxis(title = list(text = "Number of Fatalities")) |>
  hc_xAxis(title = list(text = "Year")) |>
  hc_legend(align = "center", verticalAlign = "bottom") |>
  hc_exporting(enabled = TRUE)
```
## Fatal Crash Rate

```{r, echo=FALSE}
library(highcharter)

fatal_rate <- data.frame(
  Year = c(2018, 2019, 2020, 2021, 2022, 2023, 2024),
  Rate = c(1.29, 1.21, 1.4, 1.56, 1.63, 1.22, 1.38)
)

highchart() |>
  hc_chart(type = "line") |>
  hc_title(text = "Fatal Crash Rate") |>
  hc_subtitle(text = "Palm Beach County") |>
  
  hc_xAxis(categories = fatal_rate$Year) |>
  hc_yAxis(title = list(text = "Rate per 100M VMT")) |>
  
  hc_add_series(
    name = "Fatal Crash Rate",
    data = fatal_rate$Rate,
    color = "#1f77b4"
  )
```
## Serious Injuries

```{r, echo=FALSE}
library(highcharter)

serious_injuries <- data.frame(
  Year = c(2018, 2019, 2020, 2021, 2022, 2023, 2024),
  Number = c(1194, 1020, 916, 890, 818, 766, 683)
)

hchart(serious_injuries, "column", hcaes(x = Year, y = Number)) |>
  hc_title(text = "Serious Injuries by Year") |>
  hc_subtitle(text = "Palm Beach County") |>
  hc_yAxis(title = list(text = "Number of Serious Injuries")) |>
  hc_xAxis(title = list(text = "Year"))
```
## Serious Injury Rate

```{r, echo=FALSE}
library(highcharter)

serious_rate <- data.frame(
  Year = c(2018, 2019, 2020, 2021, 2022, 2023, 2024),
  Rate = c(8.42, 7.08, 6.91, 6.45, 5.73, 5.16, NA)
)

highchart() |>
  hc_chart(type = "line") |>
  hc_title(text = "Serious Injury Rate") |>
  hc_subtitle(text = "Palm Beach County") |>
  
  hc_xAxis(categories = serious_rate$Year) |>
  hc_yAxis(title = list(text = "Rate per 100M VMT")) |>
  
  hc_add_series(
    name = "Serious Injury Rate",
    data = serious_rate$Rate,
    color = "#ff7f0e"
  )
```
# 2) Infrastructure Conditions
## Bridge Conditions
## Bridges in Good Condition

```{r, echo=FALSE}
library(highcharter)

good_condition <- data.frame(
  Year = c(2017, 2018, 2019, 2020, 2021, 2022, 2023, 2024),
  Value = c(0.877, 0.874, 0.852, 0.822, 0.821, 0.817, 0.842, 0.832)
)

highchart() |>
  hc_chart(type = "line") |>
  hc_title(text = "Bridges in Good Condition") |>
  hc_subtitle(text = "Palm Beach County") |>
  
  hc_xAxis(categories = good_condition$Year) |>
  
hc_yAxis(
  title = list(text = "Percent of Bridges"),
  min = 0.4,
  max = 1,
  tickInterval = 0.1,
  labels = list(
    formatter = JS("function() {
      return (this.value * 100).toFixed(0) + '%';
    }")
  )
) |>
  
  hc_add_series(
    name = "Good Condition",
    data = good_condition$Value,
    color = "#2ca02c"
  ) |>
  
  hc_add_series(
    name = "Target",
    data = rep(0.50, 8),
    dashStyle = "Dash",
    color = "#000000"
  ) |>
  
  hc_plotOptions(line = list(marker = list(enabled = TRUE)))
```
## Bridges in Poor Condition

```{r, echo=FALSE}
library(highcharter)

poor_condition <- data.frame(
  Year = c(2017, 2018, 2019, 2020, 2021, 2022, 2023, 2024),
  Value = c(0.012, 0.010, 0.010, 0.010, 0.010, 0.022, 0.000, 0.000)
)

highchart() |>
  hc_chart(type = "line") |>
  hc_title(text = "Bridges in Poor Condition") |>
  hc_subtitle(text = "Palm Beach County") |>
  
  hc_xAxis(categories = poor_condition$Year) |>
  
 hc_yAxis(
  title = list(text = "Percent of Bridges"),
  min = 0,
  max = 0.06,
  tickInterval = 0.01,
  labels = list(
    formatter = JS("function() {
      return (this.value * 100).toFixed(0) + '%';
    }")
  )
) |>
  
  hc_add_series(
    name = "Poor Condition",
    data = poor_condition$Value,
    color = "#d62728"
  ) |>
  
  hc_add_series(
    name = "Target",
    data = rep(0.05, 8),
    dashStyle = "Dash",
    color = "#000000"
  ) |>
  
  hc_plotOptions(line = list(marker = list(enabled = TRUE)))
```
## Map of All Bridges

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(leaflet)
library(htmltools)

bridges <- read.csv("C:/MPO/OneDrive - Palm Beach MPO/Documents/Performance_Measures/data/Bridges.csv")

bridges <- bridges[
  !is.na(bridges$Latitude) &
  !is.na(bridges$Longitude) &
  bridges$Latitude >= -90 & bridges$Latitude <= 90 &
  bridges$Longitude >= -180 & bridges$Longitude <= 180,
]

condition_order <- factor(
  bridges$Condition,
  levels = c("Good", "Fair", "Poor")
)

pal <- colorFactor(
  palette = c("green", "orange", "red"),
  levels = levels(condition_order),
  ordered = TRUE
)

labs <- paste0(
  "<b>Bridge ID:</b> ", bridges$Bridge_ID,
  "<br><b>Condition:</b> ", bridges$Condition
)

leaflet(data = bridges) |>
  addProviderTiles(providers$CartoDB.Positron) |>
  addCircleMarkers(
    lng = ~Longitude,
    lat = ~Latitude,
    popup = ~paste0(
      "<b>Bridge ID:</b> ", Bridge_ID,
      "<br><b>Condition:</b> ", Condition
    ),
    radius = 5,
    label = lapply(labs, htmltools::HTML),
    color = ~pal(Condition),
    weight = 1,
    fillOpacity = 0.8
  ) |>
  setView(lng = -80.15, lat = 26.65, zoom = 9) |>
  addLegend(
    position = "bottomright",
    pal = pal,
    values = condition_order,
    title = "Bridge Condition",
    opacity = 1
  )
```
## Pavement Conditions
### Interstate Pavements in Good Condition

```{r, echo=FALSE}
library(highcharter)

interstate_good <- data.frame(
  Year = c(2016, 2017, 2018, 2019, 2020, 2021, 2022, 2023, 2024),
  Value = c(0.624, 0.552, 0.232, 0.612, 0.532, 0.595, 0.650, 0.673, 0.547)
)

highchart() |>
  hc_chart(type = "line") |>
  hc_title(text = "Interstate Pavements in Good Condition") |>
  hc_subtitle(text = "Palm Beach County") |>
  hc_xAxis(categories = interstate_good$Year) |>
  hc_yAxis(
  title = list(text = "Percent"),
  min = 0.2,
  max = 0.8,
  tickInterval = 0.1,
  labels = list(
    formatter = JS("function() {
      return (this.value * 100).toFixed(0) + '%';
    }")
  )
) |>
  hc_add_series(
    name = "Good Condition",
    data = interstate_good$Value,
    color = "#2ca02c"
  ) |>
  hc_add_series(
    name = "Target",
    data = rep(0.60, 9),
    dashStyle = "Dash",
    color = "#888888",
    lineWidth = 1
  ) |>
  hc_plotOptions(line = list(marker = list(enabled = TRUE)))
```

### Interstate Pavements in Poor Condition

```{r, echo=FALSE}
library(highcharter)

interstate_poor <- data.frame(
  Year = c(2016, 2017, 2018, 2019, 2020, 2021, 2022, 2023, 2024),
  Value = c(0.000, 0.000, 0.000, 0.000, 0.002, 0.000, 0.000, 0.000, 0.000)
)

highchart() |>
  hc_chart(type = "line") |>
  hc_title(text = "Interstate Pavements in Poor Condition") |>
  hc_subtitle(text = "Palm Beach County") |>
  hc_xAxis(categories = interstate_poor$Year) |>
  hc_yAxis(
  title = list(text = "Percent"),
  min = 0,
  max = 0.06,
  tickInterval = 0.01,
  labels = list(
    formatter = JS("function() {
      return (this.value * 100).toFixed(0) + '%';
    }")
  )
) |>
  hc_add_series(
    name = "Poor Condition",
    data = interstate_poor$Value,
    color = "#d62728"
  ) |>
  hc_add_series(
    name = "Target",
    data = rep(0.05, 9),
    dashStyle = "Dash",
    color = "#888888",
    lineWidth = 1
  ) |>
  hc_plotOptions(line = list(marker = list(enabled = TRUE)))
```
### Non-Interstate Pavements in Good Condition

```{r, echo=FALSE}
library(highcharter)

nonint_good <- data.frame(
  Year = c(2016, 2017, 2018, 2019, 2020, 2021, 2022, 2023, 2024),
  Value = c(0.417, 0.403, 0.399, 0.440, NA, 0.451, 0.533, 0.564, 0.563)
)

highchart() |>
  hc_chart(type = "line") |>
  hc_title(text = "Non-Interstate Pavements in Good Condition") |>
  hc_subtitle(text = "Palm Beach County") |>
  
  hc_xAxis(categories = nonint_good$Year) |>
  
  hc_yAxis(
  title = list(text = "Percent"),
  min = 0.3,
  max = 0.7,
  tickInterval = 0.1,
  labels = list(
    formatter = JS("function() {
      return (this.value * 100).toFixed(0) + '%';
    }")
  )
) |>
  
  hc_add_series(
    name = "Good Condition",
    data = nonint_good$Value,
    color = "#2ca02c"
  ) |>
  
  hc_add_series(
    name = "Target",
    data = rep(0.40, 9),
    dashStyle = "Dash",
    color = "#888888",
    lineWidth = 1
  ) |>
  
  hc_plotOptions(line = list(marker = list(enabled = TRUE)))
```
### Non-Interstate Pavements in Poor Condition

```{r, echo=FALSE}
library(highcharter)

nonint_poor <- data.frame(
  Year = c(2016, 2017, 2018, 2019, 2020, 2021, 2022, 2023, 2024),
  Value = c(0.004, 0.005, 0.001, 0.001, NA, 0.012, 0.011, 0.006, 0.005)
)

highchart() |>
  hc_chart(type = "line") |>
  hc_title(text = "Non-Interstate Pavements in Poor Condition") |>
  hc_subtitle(text = "Palm Beach County") |>
  
  hc_xAxis(categories = nonint_poor$Year) |>
 hc_yAxis(
  title = list(text = "Percent"),
  min = 0,
  max = 0.06,
  tickInterval = 0.01,
  labels = list(
    formatter = JS("function() {
      return (this.value * 100).toFixed(0) + '%';
    }")
  )
) |>
  
  hc_add_series(
    name = "Poor Condition",
    data = nonint_poor$Value,
    color = "#d62728"
  ) |>
  
  hc_add_series(
    name = "Target",
    data = rep(0.05, 9),
    dashStyle = "Dash",
    color = "#888888",
    lineWidth = 1
  ) |>
  
  hc_plotOptions(line = list(marker = list(enabled = TRUE)))
```
## Pavement Conditions Map

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(leaflet)
library(sf)

pavement <- st_read(
  "C:/MPO/OneDrive - Palm Beach MPO/Documents/Performance_Measures/data/pavement.geojson",
  quiet = TRUE
)

pavement <- st_collection_extract(pavement, "LINESTRING")

pal <- colorFactor(
  palette = c("Good" = "green", "Fair" = "orange", "Poor" = "red"),
  domain = pavement$pm_pavement
)

leaflet(pavement) |>
  addProviderTiles(providers$CartoDB.Positron) |>
  addPolylines(
    color = ~pal(pm_pavement),
    weight = 4,
    opacity = 0.85,
    popup = ~paste0(
      "<b>Pavement Condition:</b> ", pm_pavement
    )
  ) |>
  setView(lng = -80.15, lat = 26.65, zoom = 9) |>
  addLegend(
    position = "bottomright",
    pal = pal,
    values = pavement$pm_pavement,
    title = "Pavement Condition",
    opacity = 1
  )
```
# 3) System Performance
## Travel Time Reliability
### Travel Time on the Interstate

```{r, echo=FALSE}
library(highcharter)

interstate_tt <- data.frame(
  Year = c(2017, 2018, 2019, 2020, 2021, 2022, 2023, 2024),
  Value = c(0.854, 0.854, 0.780, 0.936, 0.831, 0.776, 0.755, 0.623)
)

highchart() |>
  hc_chart(type = "line") |>
  hc_title(text = "Travel Time Reliability (Interstate)") |>
  hc_subtitle(text = "Palm Beach County") |>
  
  hc_xAxis(categories = interstate_tt$Year) |>
  
  hc_yAxis(
  title = list(text = "Percent of Reliable Miles"),
  min = 0.6,
  max = 1,
  tickInterval = 0.1,
  labels = list(
    formatter = JS("function() {
      return (this.value * 100).toFixed(0) + '%';
    }")
  )
) |>
  
  hc_add_series(
    name = "Reliable Miles",
    data = interstate_tt$Value,
    color = "#1f77b4"
  ) |>
  
  hc_add_series(
    name = "Target",
    data = rep(0.70, 8),
    dashStyle = "Dash",
    color = "#888888",
    lineWidth = 1
  ) |>
  
  hc_plotOptions(line = list(marker = list(enabled = TRUE)))
```
### Travel Time on Non-Interstate Roads

```{r, echo=FALSE}
library(highcharter)

nonint_tt <- data.frame(
  Year = c(2017, 2018, 2019, 2020, 2021, 2022, 2023, 2024),
  Value = c(0.910, 0.931, 0.940, 0.980, 0.968, 0.924, 0.892, 0.868)
)

highchart() |>
  hc_chart(type = "line") |>
  hc_title(text = "Travel Time Reliability (Non-Interstate)") |>
  hc_subtitle(text = "Palm Beach County") |>
  
  hc_xAxis(categories = nonint_tt$Year) |>
  
  hc_yAxis(
  title = list(text = "Percent of Reliable Miles"),
  min = 0.5,
  max = 1,
  tickInterval = 0.1,
  labels = list(
    formatter = JS("function() {
      return (this.value * 100).toFixed(0) + '%';
    }")
  )
) |>
  
  hc_add_series(
    name = "Reliable Miles",
    data = nonint_tt$Value,
    color = "#1f77b4"
  ) |>
  
  hc_add_series(
    name = "Target",
    data = rep(0.50, 8),
    dashStyle = "Dash",
    color = "#888888",
    lineWidth = 1
  ) |>
  
  hc_plotOptions(line = list(marker = list(enabled = TRUE)))
```
## Truck Travel Time Reliability

```{r, echo=FALSE}
library(highcharter)

truck_tt <- data.frame(
  Year = c(2017, 2018, 2019, 2020, 2021, 2022, 2023, 2024),
  Value = c(1.63, 1.77, 1.86, 1.66, 1.78, 1.95, 2.02, 2.09)
)

highchart() |>
  hc_chart(type = "line") |>
  hc_title(text = "Truck Travel Time Reliability") |>
  hc_subtitle(text = "Palm Beach County") |>
  hc_xAxis(categories = truck_tt$Year) |>
  hc_yAxis(
    title = list(text = "Reliability Ratio"),
    min = 1.5,
    max = 2.2
  ) |>
  hc_add_series(
    name = "Reliability Ratio",
    data = truck_tt$Value,
    color = "#1f77b4"
  ) |>
  hc_add_series(
    name = "Target",
    data = rep(2.00, 8),
    dashStyle = "Dash",
    color = "#888888",
    lineWidth = 1
  ) |>
  hc_plotOptions(line = list(marker = list(enabled = TRUE)))
```